$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-04 paper/paper.tex
Correct little slip

We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
These leftovers can then be used as an input to another typing derivation.
Leftover typing presents three main advantages:
\begin{itemize}
  \item
    The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgment with both an input and an output usage context.
  \item
    No need for an extensional context splitting proof for parallel composition.
    Alternatively, such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
    This could be obtained through automated search --- but it is better yet not to need it.
    Leftover typing uses the leftovers of $P$ to type $Q$.
  \item
    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated. \todo{explain suddenly}
\end{itemize}

We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
Both are indexed by a typing context, an input usage context, and an output usage context.

@@

\begin{example}
  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
  \begin{lstlisting}[mathescape,caption=Here the underscore \_ introduces an anonymous declaration immediately followed by its definition.]
  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
  _ = $\constr{there} \; \constr{here}$
  \end{lstlisting}
\end{example}

>>>>

We use \emph{leftover typing} \cite{Allais2018a} for our typing rules, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
These leftovers can then be used as an input to another typing derivation.
This technique presents three main advantages:
\begin{itemize}
  \item
    The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgment with both an input and an output usage context.
  \item
    The need for an extensional context splitting proof for parallel composition is eradicated.
    Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
    This could be obtained through automated search --- but it is better yet not to need it.
    Leftover typing uses the leftovers of $P$ to type $Q$.
  \item
    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated.
\end{itemize}

We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
Both of these are indexed by a typing context, an input usage context, and an output usage context.

@@

\begin{example}
  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\suc \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
  \begin{lstlisting}[mathescape,caption=Here the underscore \_ introduces an anonymous declaration immediately followed by its definition.]
  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
  _ = $\constr{there} \; \constr{here}$
  \end{lstlisting}
\end{example}
